---
title: "p8105_hw6_brm2150"
author: "Brooklynn McNeil"
date: "2024-11-19"
output: github_document
---
```{r setup, include=FALSE}
library(tidyverse)

knitr::opts_chunk$set(
  comment = '', fig.width = 8, fig.height = 6, out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```

## Problem 1

## Problem 2

Load the homicide data. Create a binary variable for the status for solved/not solved. Filter out cities that don't report race, and only include Black and White races

```{r}
url = "https://raw.githubusercontent.com/washingtonpost/data-homicides/master/homicide-data.csv"
homicide_dat = 
  read_csv(url, na = "Unknown") |>
  mutate(city_state = str_c(city, ", ", state),
         status = case_when(
           disposition %in% c("Closed without arrest", "Closed by arrest") ~ 0,
           disposition == "Open/No arrest" ~ 1
           )) |>
  filter(!city_state %in% c("Dallas, TX", "Phoenix, AZ", "Kansas City, MO", "Tulsa, AL"),
         victim_race %in% c("White", "Black"))

head(homicide_dat)
```
Create a logistic regression for `status` as the outcome and `age`, `race`, and `sex` as predictors for Baltimore, MD. 

```{r}
baltimore_glm = 
  homicide_dat |>
  filter(city_state == "Baltimore, MD") |>
  glm(status ~ victim_age + victim_race + victim_sex, data = _, family = "binomial")

tidy_results = broom::tidy(baltimore_glm, conf.int = TRUE, exponentiate = TRUE)

# show odds ratio for men
tidy_results |>
  filter(term == "victim_sexMale") |>
  select(-c(std.error, statistic, p.value))
```

Now let's map this across all of the cities, and then plot the odds ratios for murders of men being solved compared to compare across cities.

```{r}
citywide_odds = 
  homicide_dat |>
  group_by(city_state) |>
  nest() |>
  mutate(
    glm = map(data, \(df) glm(status ~ victim_age + victim_race + victim_sex, data = df, family = "binomial")),
    glm = map(glm, \(i) broom::tidy(i, conf.int = TRUE, exponentiate = TRUE)),
  ) |>
  unnest(glm) |>
  filter(term == "victim_sexMale")

citywide_odds |>
  mutate(city_state = factor(city_state, levels = citywide_odds$city_state[order(citywide_odds$estimate)])) |>
  ggplot(aes(x = city_state, y = estimate)) +
  geom_point() +
  geom_errorbar(aes(ymin = conf.low, ymax = conf.high)) +
  geom_hline(yintercept = 1, linetype = "dashed", color = "red") +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1)) +
  labs(title = "Odds Ratios for Homicides Status by Sex",
       x = ("City, State"),
       y = "odds ratio")
```
It looks like most cities have a higher ratio of homicides of men getting solved than to women. The city where the highest ratio is in New York. Long Beach, CA has a very wide confidence interval, but even so the interquantile range does not include an odds ratio that men's homcides are not solved more often. 

## Problem 3

Read in the birth weight data. 

```{r}
birthweight_df = 
  read_csv("data/birthweight.csv") |>
  mutate(across(c("babysex", "frace", "malform", "mrace"), factor))

any(is.na(birthweight_df))
```

Let's take a quick look at the data. It's normally distributed. 
```{r}
birthweight_df |>
  ggplot(aes(x = bwt)) +
  geom_histogram()
```

Let's make some linear models for each of the variables.

```{r}
birthweight_df |>
  lm(bwt ~ fincome, data = _) |>
  broom::tidy()
```

Now repeat for all predictors.

```{r}
predictors = 
  birthweight_df |> 
  select(-bwt) |>
  colnames() 

single_mod_results = 
  map(predictors, ~ {
  formula = as.formula(paste("bwt ~", .x))
  lm(formula, data = birthweight_df) |>
    broom::tidy() |>
    mutate(predictor = .x)  # Add column to identify predictor
}) |>
  bind_rows()|>
  filter(term != "(Intercept)")
```

Let's look at how all of the predictors did.

```{r}
single_mod_results |>
  ggplot(aes(x = term, y = log(p.value))) +
  geom_point()
```

